---
title: "P8105_hw2_yx2638"
author: "Yifei Xu"
date: "2022-09-28"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
```


# Problem 1

```{r, message=FALSE}
# Read and clean the data
transit_df = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
        janitor::clean_names() %>%
        select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, 
               entrance_type,  ada) %>%
        mutate(entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE"),
               entry = as.logical(entry))


# names of important variables
variable_name = names(transit_df)

# size of the dataset 
row = nrow(transit_df)
col = ncol(transit_df)
```

```{r, message=FALSE}
# Count the number of distinct stations
station_num = nrow(distinct(transit_df, line, station_name))

# Count the number of stations that are ADA compliant
ada_station_num = transit_df %>%
        filter(ada == "TRUE") %>%
        distinct(line, station_name) %>%
        nrow()

# Calculte the proportion of station entrances / exits without vending allow entrance ???distinct
no_vending_extrance_num = transit_df %>%
        filter(entry == "TRUE", vending == "NO") %>%
        nrow()

no_vending_num = transit_df %>%
        filter(vending == "NO") %>%
        nrow()

prop = scales::percent(no_vending_extrance_num / no_vending_num, 0.01)
```

There are `r station_num` distinct stations and `r ada_station_num` stations are ADA compliant.
The proportion of station entrances / exits without vending allow entrance is `r prop`.


```{r, message=FALSE}
# Reformat data so that route number and route name are distinct variables
transit_df_tidy = transit_df %>%
        mutate(
                route8 = as.character(route8),
                route9 = as.character(route9),
                route10 = as.character(route10),
                route11 = as.character(route11)
               ) %>%
        pivot_longer(
                route1:route11,
                names_to = "route_number",
                values_to = "route_name"
                )

# Count the number of distinct stations serve the A train
station_a_num = transit_df_tidy %>%
        filter(route_name == "A") %>%
        distinct(line, station_name) %>%
        nrow()

# Count the number of ADA compliant of the stations that serve the A train
station_a_ada_num = transit_df_tidy %>%
        filter(route_name == "A", ada == "TRUE") %>%
        distinct(line, station_name) %>%
        nrow()
```

`r station_a_num` distinct stations serve the A train, of which `r station_a_ada_num` are ADA compliant.


# Problem 2

```{r, message=FALSE}
# Read and clean the Mr. Trash Wheel sheet
mr_trash = readxl::read_excel("data/Trash Wheel Collection Data.xlsx", 
                              sheet = "Mr. Trash Wheel", 
                              range = cellranger::cell_cols("A:N")
                              )

mr_trash_tidy = mr_trash %>%
        janitor::clean_names() %>%
        drop_na(dumpster) %>%
        mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))
```

```{r, message=FALSE}
# Read and clean the Professor Trash Wheel sheet
prof_trash = readxl::read_excel("data/Trash Wheel Collection Data.xlsx", 
                                sheet = "Professor Trash Wheel", 
                                range = cellranger::cell_cols("A:M") 
                                )

prof_trash_tidy = prof_trash %>%
        janitor::clean_names() %>%
        drop_na(dumpster) 

```

```{r, message=FALSE}
# Add a new variable in these two datasets
mr_trash_tidy_comb = mr_trash_tidy %>%
        mutate(wheel_name = "Mr. Trash Wheel")

prof_trash_tidy_comb = prof_trash_tidy %>%
        mutate(wheel_name = "Professor Trash Wheel",
               sports_balls = NA) %>%
        relocate(sports_balls, .before = "homes_powered") 

# Combine these two datasets
mr_prof_trash = rbind(mr_trash_tidy_comb, prof_trash_tidy_comb) %>%
        select(wheel_name, everything())

```

```{r, message=FALSE}
# Count the number of observations in the resulting dataset
mr_prof_nrow = nrow(mr_prof_trash)

# Give examples of key variables
mr_prof_var = names(mr_prof_trash)

# Calculate the total weight of trash collected by Professor Trash Wheel
prof_total_weight = mr_prof_trash %>%
        filter(wheel_name == "Professor Trash Wheel") %>%
        pull(weight_tons) %>%
        sum()

# Count the total number of sports balls collected by Mr. Trash Wheel in 2020
mr_total_balls = mr_prof_trash %>%
        filter(wheel_name == "Mr. Trash Wheel", year == "2020") %>%
        pull(sports_balls) %>%
        sum()
```
* `Mr. Trash Wheel` dataset  
The size of the tidy dataset is `r nrow(mr_trash_tidy)` rows * `r ncol(mr_trash_tidy)` columns.  
The key variables include `r names(mr_trash_tidy)`.  

* `Professor Trash Wheel` dataset  
The size of the tidy dataset is `r nrow(prof_trash_tidy)` rows * `r ncol(prof_trash_tidy)` columns.  
The key variables include `r names(prof_trash_tidy)`.  

* Resulting(combining) dataset  
There are `r mr_prof_nrow` observations in the resulting dataset and the key variables include `r mr_prof_var`.  
The total weight of trash collected by Professor Trash Wheel was `r prof_total_weight` tons.  
The total number of sports balls collected by Mr. Trash Wheel in 2020 was `r mr_total_balls`.


# Problem 3

```{r, message=FALSE}
# Read and clean the data in pols-month.csv
pols = read_csv("data/pols-month.csv")

pols_tidy = pols %>%
        janitor::clean_names() %>%
        separate(mon, into = c("year", "month", "day")) %>%
        mutate(
                month = month.name[as.numeric(month)],
                year = as.numeric(year),
                president = case_when(prez_dem == 1 ~ "dem",
                               prez_gop == 1 ~ "gop",
                               prez_gop == 2 ~ "gop") 
                )%>%
         arrange(year, month) %>%
        select(-day, -prez_dem, -prez_gop)
```

```{r, message=FALSE}
# Read and clean the data in snp.csv
snp = read_csv("data/snp.csv")

snp_tidy = snp %>%
        janitor::clean_names() %>%
        separate(date, into = c("month", "day", "year"), sep ="/") %>%
        mutate(
                month = month.name[as.numeric(month)],
                year = as.numeric(year),
                year = ifelse(year < 23, year+2000, year+1900)
                ) %>%
        arrange(year, month) %>%
        select(year, month, close)
```

```{r, message=FALSE}
# Read and clean the data in unemployment.csv
unemployment = read_csv("data/unemployment.csv")

unemployment_tidy = unemployment %>%
        pivot_longer(
                Jan:Dec,
                names_to = "month",
                values_to = "unemployment_percent"
        ) %>%
        janitor::clean_names() %>% 
        mutate(
                month = match(month, month.abb),
                month = month.name[month]
                ) %>%
        arrange(year, month) 

```

```{r, message=FALSE}
# Join the datasets by merging snp into pols
pols_snp = left_join(pols_tidy, snp_tidy, by = c("year","month"))

# Join the datasets by merging unemployment into the result
pols_snp_employment = left_join(pols_snp, unemployment_tidy, by = c("year","month"))
```

```{r, message=FALSE}
# the dimension of the resulting dataset 
res_nrow = nrow(pols_snp_employment)
res_ncol = ncol(pols_snp_employment)

# range of years
res_year = pols_snp_employment %>%
        pull(year) %>% 
        range()

res_year_min = pols_snp_employment %>%
        pull(year) %>% 
        min()

res_year_max = pols_snp_employment %>%
        pull(year) %>% 
        max()

# names of key variables
res_var = names(pols_snp_employment)
```
* `pols-month` dataset  
The tidy `pols-month` dataset contains `r nrow(pols_tidy)` observations and `r ncol(pols_tidy)` variables related to the number of national politicians who are democratic or republican from `r pols_tidy %>% pull(year) %>% min` to `r pols_tidy %>% pull(year) %>% max`. Key variables include `r names(pols_tidy)`

* `snp` dataset  
The tidy `snp` dataset contains `r nrow(snp_tidy)` observations and `r ncol(snp_tidy)` variables related to Standard & Poor’s stock market index (S&P) with a year range from `r snp_tidy %>% pull(year) %>% min` to `r snp_tidy %>% pull(year) %>% max`. Key variables include `r names(snp_tidy)`

* `unemployment` dataset  
The tidy `unemployment` dataset contains `r nrow(unemployment_tidy)` observations and `r ncol(unemployment_tidy)` variables related to the number of national politicians who are democratic or republican from `r unemployment_tidy %>% pull(year) %>% min` to `r unemployment_tidy %>% pull(year) %>% max`. Key variables include `r names(unemployment_tidy)`

* Merged dataset `pols_snp_unemploymet`  
The merged dataset `pols_snp` contains `r res_nrow` observations and `r res_ncol` variables, with a range of years from `r res_year_min` to `r res_year_max`. Key variables includes `r res_var`, which represents year and month of the observation, the number of republican governors, the number of republican senators, the number of republican representatives, the number of democratic governors, the number of democratic senators, the number of democratic representatives on the associated date, the president's political party, the closing values of the S&P stock index and percentage of unemployment on the associated date.By joining these three datasets, we can further analze how president's political party affect economic outcomes such as stocks and unemployment.








